---
title: "Employee Churn- Survival Analysis"
author: "Karol Orozco"
date: "2022-10-04"
---

## Import Libraries

```{r, echo=FALSE}
library(tidyverse)
library(ggplot2)
library(survival)
library(skimr)

```

## Load the Data

This data set came from Kaggle

```{r}
employee.att <- read.csv("C:/Users/karol/Desktop/Portfolio/DATASET/IBM-HR-Employee-Attrition.csv")
```

Let's take a look at the data:

```{r}
skim(employee.att)
```

This data set contains 1470 observations and 35 variables. However, for this analysis, I am interested in the following fields:

-   Attrition:

-   Department:

-   Age:

-   Gender:

-   EnvironmentSatisfaction:

-   JobLevel: JobSatisfaction:

-   YearsAtCompany:

-   YearsSinceLastPromotion:

-   OverTime:

## Survival Analysis

```{r}

#as 'Attrition' is a categorical variable, the following command changes it to a binary 1/0

employee.att$event <- with(employee.att,ifelse(Attrition=="Yes", 1, 0))

time<-employee.att$YearsAtCompany #define time

event<-employee.att$event #define event

group<-employee.att$OverTime #define groups to compare multiple survival curves

survival<-Surv(time,event) #creates the survival object

head(survival, n = 15)
```
