---
title: "Time Series Project"
author: "Karol Orozco"
date: "2022-10-04"
categories: [news, code, analysis]
---

```{r, message=FALSE, warning=FALSE}

library(tidyverse)
library(lubridate); library(tsibble)
library(readxl)
library(purrr)
library(gganimate)
library(fpp3)

```

## Load the Data

You can find this free dataset on the Zillow website: https://www.zillow.com/research/data/#other-metrics


```{r}
raw_city <- read.csv("C:/Users/karol/Desktop/Portfolio/City_zori_sm_month.csv")
```

- RegionID	This is unique Id for the Regions
- SizeRank	This is the ranking done based on the size of the region
- RegionName	This field contains the zip code of the region.
- RegionType	Type of region is Zip.
- StateName	State
- State	This column provide the specific City Name of Housing Data
- Metro	This provide the name of the metro city around that region
- County Name	This is the county name for that region
- Months Column	These columns contains the prices of region for every month since 2015- 2022

```{r}

str(raw_city[,c(1:11)])
```

We have to make this dataset tidy. Tidy Data is a way of structuring data so that it can be easily understood by people and analyzed by machines.

## Prepare and Convert the Data

I need to remove the X at the beginning of the dates (X3.31.2015,...)

```{r}
names(raw_city) <- sub("^X", "", names(raw_city))
str(raw_city[,c(1:10)],10)
```

Now, I will create a new column called Price, where the values from the dates column will merge into it.

```{r}
rent_city <- raw_city %>% 
  pivot_longer(-c(RegionID,SizeRank,RegionName,RegionType,StateName,State,Metro,CountyName),
    names_to = "Monthly",
    values_to = "Price"
  ) 
str(rent_city)

```




```{r}
#Converting the Date from factor to character
rent_clean <- rent_city %>%
            mutate(Monthly_parsed = as.Date(Monthly,"%m.%d.%Y"))


rent_clean[["Monthly"]]<- as.character(rent_clean$Monthly)

rent_city[["Monthly"]]<- as.character(rent_city$Monthly)
summary(rent_city)
```


We see some missing values in the Price variable, but before I deal with those values, I will filter my data to the cities that I am interested the most: Houston, Orlando and Portland.


```{r}

preferred_cities <- rent_clean %>%
  filter(rent_clean$RegionName %in% c("Houston", "Orlando", "Portland"))

summary(preferred_cities)
```


After filtering the data, we don't have any missing values

```{r}
colSums(is.na(preferred_cities))
```

```{r}
ts_rent_pref_cities <- preferred_cities[,c("Monthly_parsed","Price")]


#### Display the first 10 elements of the dataset

head(ts_rent_pref_cities, n = 10)

#### Display the last 10 elements of the dataset

tail(ts_rent_pref_cities, n = 10)

```
We see that the date goes from March 31, 2015, to September 30, 2022, with a monthly calculation of the rental value.

```{r, fig.width= 15, message=FALSE, warning=FALSE}

preferred_cities %>%
  ggplot(aes(x= Monthly_parsed, y= Price, color= RegionName)) +
  geom_line(size=2) +
  labs(y="Rent Price in /n US Dollars", 
       x= "Year",
       title="Rent Cost in Portland, Orlando & Houston, 2015-2022",
       caption = "data:https://www.zillow.com/research/data")+
  facet_grid(RegionName~ ., scales = "free_y")+
  hrbrthemes::theme_ipsum_rc()+  
  theme(legend.position = "none") 

```

# Decomposition

## Classical Decomposition




